Quantifying the home ground advantage

### 2022-04-20 inst/home_advantage.Rmd
use sticker in Rmd
label each chunk, and use cache

@@

First load the package:
```{r}
library(hyper2)
```

@@

```{r}
home_games_won <- matrix(c(
    NA, 16, 12, 11,
    19, NA, 19, 16,
    17, 12, NA, 11,
    11, 12, 12, NA),
    nrow=4,ncol=4,byrow=TRUE)

@@

```{r}
H <- hyper2(pnames=c(teams,"home"))
```

@@

```{r}
for(i in seq_len(nrow(home_games_won))){
  for(j in seq_len(ncol(home_games_won))){
    if(i != j){  # diagonal means a team plays itself, meaningless
      game_winner <- teams[i]
      game_loser  <- teams[j]
      no_of_matches <- home_games_won[i,j] # won with home strength helping
      H[c(game_winner,           "home")] %<>% inc(no_of_matches)  # won with home strength helping
      H[c(game_winner,game_loser,"home")] %<>% dec(no_of_matches)
    }
  }
}
```

@@

```{r}
for(i in seq_len(nrow(away_games_won))){
  for(j in seq_len(ncol(away_games_won))){
    if(i != j){
      game_winner <- teams[i]
      game_loser  <- teams[j]
      no_of_matches <- away_games_won[i,j] # won without the benefit of home strength
      H[c(game_winner                  )] %<>% inc(no_of_matches)  # won without home ground helping...
      H[c(game_winner,game_loser,"home")] %<>% dec(no_of_matches)  # home strength nevertheless on denominator
    }
  }
}
```

@@

```{r}
H
```

@@

```{r}
dotchart(maxp(H))
```

@@

```{r}
f <- function(home){
	small <- 1e-3
	maxp(H,startp=rep((1-small)/4,4),give=TRUE,fcm=t(rep(1,4)),fcv=1-home)$value
}
```

@@

```{r}
Lambda <- -2*(f(0.01)-maxp(H,give=TRUE)$value)
Lambda
```

@@

```{r}
g <- function(home){
	small <- 1e-3
	maxp(H,startp=rep(small/4,4),give=TRUE,fcm=-t(rep(1,4)),fcv=home-1)$value
}
```

```{r}
home_strength <- seq(from=0.2,to=0.6,length=30)
support <- pmin(sapply(home_strength,f),sapply(home_strength,g))
plot(home_strength,support-max(support),
     xlab="home strength",ylab="support",main="profile likelihood for home advantage")
abline(h=c(0,-2))
```

>>>>

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

(takes about ten minutes to run without cache)

@@

First load the package:
```{r,label=libload}
library(hyper2)
```

@@

```{r,label=genthedata}
home_games_won <- matrix(c(
    NA, 16, 12, 11,
    19, NA, 19, 16,
    17, 12, NA, 11,
    11, 12, 12, NA),
    nrow=4,ncol=4,byrow=TRUE)

@@

```{r,label=makehyper2}
H <- hyper2(pnames=c(teams,"home"))
```

@@

```{r,label=includehome}
for(i in seq_len(nrow(home_games_won))){
  for(j in seq_len(ncol(home_games_won))){
    if(i != j){  # diagonal means a team plays itself, meaningless
      game_winner <- teams[i]
      game_loser  <- teams[j]
      no_of_matches <- home_games_won[i,j] # won with home strength helping
      H[c(game_winner,           "home")] %<>% inc(no_of_matches)  # won with home strength helping
      H[c(game_winner,game_loser,"home")] %<>% dec(no_of_matches)
    }
  }
}
```

@@

```{r,label=awaygameswon}
for(i in seq_len(nrow(away_games_won))){
  for(j in seq_len(ncol(away_games_won))){
    if(i != j){
      game_winner <- teams[i]
      game_loser  <- teams[j]
      no_of_matches <- away_games_won[i,j] # won without the benefit of home strength
      H[c(game_winner                  )] %<>% inc(no_of_matches)  # won without home ground helping...
      H[c(game_winner,game_loser,"home")] %<>% dec(no_of_matches)  # home strength nevertheless on denominator
    }
  }
}
```

@@

```{r,label=showH}
H
```

@@

```{r,label=showvisH}
dotchart(maxp(H))
```

@@

```{r,label=nullofzerohomestrength}
f <- function(home){
	small <- 1e-3
	maxp(H,startp=rep((1-small)/4,4),give=TRUE,fcm=t(rep(1,4)),fcv=1-home)$value
}
```

@@

```{r,label=showLambda,cache=TRUE}
Lambda <- -2*(f(0.01)-maxp(H,give=TRUE)$value)
Lambda
```

@@

```{r,calcprofilelikelihood}
g <- function(home){
	small <- 1e-3
	maxp(H,startp=rep(small/4,4),give=TRUE,fcm=-t(rep(1,4)),fcv=home-1)$value
}
```

```{r,showprofilelikelihood,cache=TRUE}  # this one takes the time
home_strength <- seq(from=0.2,to=0.6,length=30)
support <- pmin(sapply(home_strength,f),sapply(home_strength,g))
plot(home_strength,support-max(support),
     xlab="home strength",ylab="support",main="profile likelihood for home advantage")
abline(h=c(0,-2))
```
